---
title: "annulus_demo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{annulus_demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ashapesampler)
library(alphahull)
library(ggplot2)
library(doparallel)
cores = detectCores()
registerDoParallel(cores=cores)
```

In this document, we demonstrate the $\alpha$-shape sampler pipeline by simulating
the process of learning a set of two-dimensional shapes (in this case, annuli) and
simulating a new shape from that. This vignette requires the packages \emph{alphahull} and \emph{ggplot2} in addition to \emph{ashapesampler}.

We begin by setting the parameters for our simulation. We will fix $\alpha=0.25$ 
and $n=100$, and draw 20 shapes for our data set. Our true underlying manifold
will be the annulus with outer radius 0.75 and inner radius 0.25.

```{r}
set.seed(201723)
my_alpha = 0.25
n = 100
N=20
r_maj <- 0.75
r_min <- 0.25
```

Next we will draw the shapes themselves.

```{r}
ann_list <- list()
complex_list <- list()
tau_vec <- vector("numeric", N)
for (k in 1:N){
  ann_pts <- runif_annulus(n, r_maj, r_min)
  ann_list[[k]] <- ashape(ann_pts, alpha = my_alpha)
  complex_list[[k]] <- get_alpha_complex(ann_pts, my_alpha)
  tau_vec[k] <- tau_bound(ann_list[[k]]$x, complex_list[[k]])
}
```

Now that we have the shapes generated and imported, we can learn the underlying
manifold. First we will put together the entire point cloud into one two column 
matrix.

```{r}
point_cloud = matrix(NA,nrow=0, ncol = 2)
nbr_list = c()
for (k in 1:N) {
  point_cloud = rbind(point_cloud, ann_list[[k]]$x)
}
```

Then we will have our $\tau$ bound be a summary statistic of the $\tau$ found for 
each input shape. Here, we will use mean, but one can tweak this to see different
results. Note that if $\tau$ is too small, then the random walk won't be able to 
execute around the point cloud, but if $\tau$ is too big, then we risk losing 
geometric and topological information in the reconstruction.

```{r}
tau = min(tau_vec)
```

Now we can take the parameters and generate a new shape and plot it. We assume
our bound to be a square, though the function allows for "circle" (disk) and "annulus" 
as well.
```{r}
new_annulus <- generate_ashape2d(point_cloud, N=N, tau=tau, cores=cores)


tri_keep = new_annulus$delvor.obj$tri.obj$trlist[which(new_annulus$delvor.obj$tri.obj$cclist[,3]<new_annulus$alpha), 1:3]
dim_tri = dim(tri_keep)[1]
tri_keep = as.vector(t(tri_keep))
triangles = data.frame("id"=sort(rep(1:dim_tri, 3)), "x"=new_annulus$x[tri_keep, 1], "y"=new_annulus$x[tri_keep,2])
extremes = as.data.frame(new_annulus$x[new_annulus$alpha.extremes,])

edges = as.data.frame(new_annulus$edges[,3:6])

ggplot(data.frame(new_annulus$x), aes(x=X1, y=X2)) +
  geom_polygon(data=triangles, aes(x=x, y=y, group=id), fill="gray") +
  geom_segment(data=edges, aes(x=x1, y=y1, xend=x2, yend=y2), color="blue")+
  geom_point(data=extremes, aes(x=V1, y=V2), size=1.5)+
  theme_classic()
```


