---
title: "torus_demo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{torus_demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ashapesampler)
library(alphashape3d)
library(alphahull)
library(doParallel)
cores = detectCores()
registerDoParallel(cores=cores)
```

In this document, we demonstrate the $\alpha$-shape sampler pipeline by simulating
the process of learning a set of three-dimensional shapes (in this case, tori) and
simulating a new shape from that. This vignette requires the packages \emph{alphashape3d}, \emph{alphahull}, and \emph{rgl} in addition to \emph{ashapesampler}.

We begin by setting the parameters for our simulation. We will fix $\alpha=0.25$ 
and $n=500$, and draw 20 shapes for our data set. Our true underlying manifold
will be the torus with major radius 0.75 and minor radius 0.25.

```{r}
N=10
tau_vec <- vector("numeric", N)
my_alpha <- 0.25
n <- 5000
r_maj <- 0.75
r_min <- 0.25
```

Next we will draw the shapes themselves.

```{r}
torus_list <- list()
complex_torus_list <- list()

for (k in 1:N){
  torus_pts <- rtorus(n, r_min, r_maj)
  temp_torus <- ashape3d(torus_pts, my_alpha)
  torus_list[[k]] <- temp_torus
  complex_torus_list[[k]] <- get_alpha_complex(torus_pts, my_alpha)
  tau_vec[k] <- tau_bound(torus_list[[k]]$x, complex_torus_list[[k]])
}
```

Now that we have the shapes generated and imported, we can learn the underlying
manifold. First we will put together the entire point cloud into one two column 
matrix.

```{r}
choose_2 <- sample(N,2)
point_cloud = rbind(torus_list[[choose_2[1]]]$x, torus_list[[choose_2[[2]]]]$x)
```

Then we will have our $\tau$ bound be a summary statistic of the $\tau$ found for 
each input shape. Here, we will use mean, but one can tweak this to see different
results. Note that if $\tau$ is too small, then the random walk won't be able to 
execute around the point cloud, but if $\tau$ is too big, then we risk losing 
geometric and topological information in the reconstruction.

```{r}
tau = min(tau_vec[choose_2[1]], tau_vec[choose_2[2]])
```

Now we can take the parameters and generate a new shape and plot it. We assume
our bound to be a cube, though the function allows for "sphere" (three-dimensional ball) and "shell" as well.

```{r}
new_torus <- generate_ashape3d(point_cloud, J=2, tau=tau, cores=cores)
plot(new_torus, indexAlpha="all")
```
